---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simFrB

### Why this simulation Tool?

Simulated omics data often serves as a basis to evaluate or create analysis tools, plan experiments and get a deeper understanding of the data. Simulations usually make many assumptions about the distribution of the data and often times simplify complex relationships observed in experimental data.

One major challenge in simulating omics data is simulating missing value patterns. Simulations might take data dependencies into consideration by including values missing non at random, but we found that such approaches still fail to reproduce results observed in experimental data.

While trying to validate results of statistical analysis tools used in Fröhlich et. al (), we observed a strong influence of the missing value pattern on statistical results if data was prepossessed with imputation, as done in the SAM method. While for experimental data, imputation strongly increased the pAUC of all statistical tests, it decreased the pAUC for simulated data.

This divergence was found to be caused by **feature dependent correlations of the missing value pattern**, that got lost in other simulation approaches. The implemented simulation tool leverages allows to leverage benchmark data in order to simulate new count matrices with realistic missing value patterns, by estimating coefficients for both intensity (linear model) and missingness (logit model) for each feature and drawing jointly to simulate data that keeps correlations found in the experimental data.

## Usage

For now this package can be installed from this github repository

```{r}
library(remotes)
remotes::install_github("kreutz-lab/simFrB")
```

Generally there or two modes of operation:

1.  Use Available Data to estimate coefficients from. This will be most accurate for Benchmark data with known ground truth.

```{r}
#exp.data <- load("pathToExpData")
#DE_idx <- grep("DE",rownames(exp.data)) #This needs to be adapted depending on how you know which feature is differentially expressed


```

1.  Use Pre calculated Coefficients stored in data/jointCoefs.rds
